---
title: "Prepare (meta)data for Movebank upload"
author: "Sanne Govaert"
date: "`r Sys.Date()`"
output:
  html_document
---

This script prepares [Ornitela](https://www.ornitela.com/), [Interrex](http://interrex-tracking.com/) and [Druid](https://druid.tech/) bird tracking data for upload to [Movebank](https://www.movebank.org/). For a project, it downloads **reference data** from a Google spreadsheet where these data are maintained and transforms these to the [Movebank data format](https://www.movebank.org/node/2381) using `dplyr::mutate()` (see below). **GPS data** are automatically synced with Movebank (see `Live Feeds` in a Movebank study). This is not yet implemented for **acceleration data** which are currently not uploaded.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
```

Set user-defined values:

```{r}
project_id <- "TURTUR_VLAANDEREN"
```

## Get tag and bird information from spreadsheet

```{r}
metadata_ornitela <- readr::read_csv(
    "https://docs.google.com/spreadsheets/d/e/2PACX-1vSU838D5jOOcPSyL21iheYdC1hEwYaR88x9qhJNLOZtmaTuK6hv23Vpz-o1b5erXur4Vw2g9Z2mQchh/pub?gid=0&single=true&output=csv",
    col_types = cols(.default = col_character())
)
metadata_interrex <- readr::read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSU838D5jOOcPSyL21iheYdC1hEwYaR88x9qhJNLOZtmaTuK6hv23Vpz-o1b5erXur4Vw2g9Z2mQchh/pub?gid=668940&single=true&output=csv",
  col_types = cols(.default = col_character())
  )
metadata_druid <- readr::read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSU838D5jOOcPSyL21iheYdC1hEwYaR88x9qhJNLOZtmaTuK6hv23Vpz-o1b5erXur4Vw2g9Z2mQchh/pub?gid=2022063817&single=true&output=csv",
  col_types = cols(
    .default = col_character())
  )
metadata_gm <- readr::read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSU838D5jOOcPSyL21iheYdC1hEwYaR88x9qhJNLOZtmaTuK6hv23Vpz-o1b5erXur4Vw2g9Z2mQchh/pub?gid=1432656981&single=true&output=csv",
  col_types = cols(.default = col_character())
  )

metadata <-
  metadata_ornitela |>
  dplyr::bind_rows(metadata_interrex) |>
  dplyr::bind_rows(metadata_druid) |>
  dplyr::bind_rows(metadata_gm)

metadata_colnames <- colnames(metadata)
```

## Get reference data (metadata)

For a project, get data on animals, tags and deployments in the [Movebank reference data format](https://www.movebank.org/node/27).

Map data:

```{r get_ref_data}
movebank_ref_data <-
  metadata |>
  dplyr::filter(project == project_id) |>
  dplyr::mutate(
    `animal-id` = dplyr::coalesce(metal_ring, colour_ring),
    `animal-nickname` = dplyr::case_when(
      str_detect(animal_name, "OT-") ~ NA_character_, # Exclude names that are the default tag name
      TRUE ~ animal_name
    ),
    `animal-ring-id` = ifelse(tag_manufacturer %in% c("Druid", "Global Messenger"), metal_ring, colour_ring),
    `animal-sex` = dplyr::recode(sex,
      "F" = "f",
      "M" = "m",
      "X" = "u",
      .missing = "u"
    ),
    `animal-taxon` = scientific_name,
    `alt-project-id` = project_id,
    `animal-life-stage` = dplyr::recode(age,
      "A" = "adult",
      "J" = "juvenile"
    ),
    `animal-mass` = animal_weight,
    `animal-mortality-type` = mortality_type,
    `animal-mortality-date` = dplyr::if_else(
      !is.na(mortality_type),
      format(
        as.POSIXct(track_session_end_date, tz = "UTC", format = "%Y-%m-%d %H:%M:%S"),
        "%Y-%m-%dT%H:%M:%SZ"
        ),
      NA_character_
    ),
    `animal-death-comments` =
      dplyr::if_else(!is.na(mortality_type), track_session_remarks, NA_character_),
    `attachment-type` = "harness",
    `deploy-off-timestamp` = format(
      as.POSIXct(track_session_end_date, tz = "UTC", format = "%Y-%m-%d %H:%M:%S"),
      "%Y-%m-%dT%H:%M:%SZ"
    ),
    `deploy-on-latitude` = release_latitude,
    `deploy-on-longitude` = release_longitude,
    `deploy-on-measurements` = animal_measurements,
    `deploy-on-timestamp` = format(
      as.POSIXct(track_session_start_date, tz = "UTC", format = "%Y-%m-%d %H:%M:%S"),
      "%Y-%m-%dT%H:%M:%SZ"
    ),
    `deployment-comments` = track_session_remarks,
    `deployment-end-type` = dplyr::case_when(
      stringr::str_detect(tolower(track_session_remarks), "dead") ~ "dead",
      stringr::str_detect(tolower(track_session_remarks), "died") ~ "dead",
      stringr::str_detect(tolower(track_session_remarks), "predated") ~ "dead",
      stringr::str_detect(tolower(track_session_remarks), "victim") ~ "dead",
      stringr::str_detect(tolower(track_session_remarks), "defect") ~ "equipment failure",
      stringr::str_detect(tolower(track_session_remarks), "malfunction") ~ "equipment failure",
      TRUE ~ NA_character_
    ),
    `manipulation-type` = dplyr::case_when(
      stringr::str_detect(manipulation_type, "manipulated other") ~ "manipulated other",
      # Birds that were hatched from egg or raised as chicks
      # e.g. placed in controlled environment and subjected to behavioural studies
      TRUE ~ "none"
    ),
    `study-site` = release_location,
    `tag-model` = ifelse(tag_manufacturer == "Druid", tag_model, NA_character_),
    `tag-readout-method` = ifelse(tag_manufacturer == "Druid", "Wi-Fi/Bluetooth", "phone network"),
    `sensor-type` = ifelse(tag_manufacturer == "Druid", "GNSS", "GPS"),
    `tag-id` = serial_number,
    `tag-manufacturer-name` = tag_manufacturer,
    `tag-mass` = tag_weight,
    `tag-serial-no` = serial_number,
    .keep = "none"
  ) |>
  dplyr::filter(!is.na(`animal-id`)) |> 
  dplyr::arrange(`tag-id`, `animal-id`)
```

Save to CSV:

```{r}
readr::write_csv(movebank_ref_data, file = here::here("data", "processed", project_id, "movebank_ref_data.csv"), na = "")
```

## Get gps data (detections)

Done via live feed.

## Get acceleration data

Not applicable.
