---
title: "Prepare CR-Birding (meta)data for Movebank upload"
author: "Sanne Govaert, Peter Desmet"
date: "`r Sys.Date()`"
output:
  html_document
---

This script prepares [CR-Birding](https://submit.cr-birding.org/) bird ringing data for upload to [Movebank](https://www.movebank.org/). For a project, it downloads **reference and observation data** from a the CR-Birding Submit website and transforms these to the [Movebank data format](https://www.movebank.org/node/2381) using `dplyr::mutate()` (see below).

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
```

Set user-defined values:

```{r}
project_id <- "LARGE_GULLS"
start_date <- "1980-01-01"
end_date <- "2024-12-31"
```

## Get CR-Birding observation data

Load custom function to create download URLs:

```{r}
source(here::here("src", "functions", "download_urls.R"))
```

Create URLs by year:

```{r}
urls <- download_urls(start_date, end_date, project_id)
```


```


## Create reference data (metadata)

For a project, get data on animals, rings and deployments in the [Movebank reference data format](https://www.movebank.org/node/27).

Map data:

```{r get_ref_data}
movebank_ref_data <-
  data
```

Save to CSV:

```{r}
readr::write_csv(movebank_ref_data, file = here::here("data", "processed", project_id, "movebank_ref_data.csv"), na = "")
```

## Get location data (observations)

Map data:

```{r get_ref_data}
movebank_location_data <-
  data
```

Save to CSV:

```{r}
readr::write_csv(movebank_location_data, file = here::here("data", "processed", project_id, "movebank_location_data.csv"), na = "")
```
