---
title: "Download UvA-BiTS metadata for a project"
author: "Peter Desmet"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(DBI)
library(here)
library(glue)
```

## Connect to UvA-BiTS database

Get connection settings from `config.yml` (not included in this repo) and connect to database:

```{r connect_to_uvabits}
uvabits <- config::get("uvabits")
con = dbConnect(
  RPostgres::Postgres(),
  host = uvabits$server,
  port = uvabits$port,
  user = uvabits$username,
  password = uvabits$password,
  dbname = uvabits$database
)

# The alternative is getting the connection defined in odbc.ini:
# con <- dbConnect(odbc::odbc(), "uvabits")
```

## Get metadata from all projects

Get data on individuals and their associated track sessions from all our bird tracking projects combined.

Query data:

```{r}
# The SQL below doesn't have any {variables}, but I opted to use glue anyway
sql <- glue_sql(read_file(here("sql", "individuals.sql")), .con = con)
individuals <- dbGetQuery(con, sql)
```

Save to csv:

```{r}
write_csv(individuals, path = here("data", "interim", "individuals.csv"), na = "")
```

## Get metadata in Movebank format for single project

For a project, get data on animals, tags and deployments in the [Movebank reference data format](https://www.movebank.org/node/27). Some values have to be defined by the user. See [here](https://www.movebank.org/node/2381#metadata) for information on all reference data terms.

Set values:

```{r}
project <- "MH_WATERLAND"
animal_life_stage <- "adult"
attachment_type <- "harness"
manipulation_type <- "none"
bird_remarks_is_nickname <- TRUE
```

Query data:

```{r}
movebank_sql <- glue_sql(read_file(here("sql", "movebank_reference_data.sql")), .con = con)
movebank_reference_data <- dbGetQuery(con, movebank_sql)
```

Save to CSV:

```{r}
write_csv(movebank_reference_data, path = here("data", "processed", project, "movebank_reference_data.csv"), na = "")
```
