---
title: "Download data from UvA-BiTS"
author:
- Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Set locale (so we use UTF-8 character encoding):

```{r}
# This works on Mac OS X, might not work on other OS
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Load libraries:

```{r}
library(tidyverse)
library(RPostgreSQL)
library(glue)
```

Set file paths (all paths should be relative to this script):
 
```{r}
# SQL
individuals_sql_file = "../sql/individuals.sql"
individual_detections_sql_file = "../sql/individual_detections.sql"

# Data
individuals_file = "../data/interim/individuals.csv"
```

# Connect to UvA-BiTS database

Get credentials from config file (not included in this repo) and create a database connection:

```{r connect_to_uvabits}
uvabits <- config::get("uvabits")
con = dbConnect(
  dbDriver("PostgreSQL"),
  user = uvabits$username,
  password = uvabits$password,
  host = uvabits$server,
  port = uvabits$port,
  dbname = uvabits$database
)
```

# Download individuals

Get data on individuals and their associated track sessions, from all our bird tracking projects combined.

Load SQL:

```{r}
# The SQL below doesn't have any {variables}, but I opted to use glue anyway
individuals_sql <- glue_sql(read_file(individuals_sql_file), .con = con)
```

Run SQL:

```{r query_individuals}
individuals <- dbGetQuery(con, individuals_sql)
```

Preview data:

```{r}
individuals %>% head()
```

Save to CSV:

```{r}
write.csv(individuals, file = individuals_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```

# Download WMH detections

Define dataset properties:

```{r}
project_keys = c("MH_WATERLAND")
download_directory = "../data/interim/wmh"
```

Get `device_info_serial`s for that project(s). Note that because of the reuse of tags, there might be less `device_info_serials` than bird individuals:

```{r}
device_info_serials <- individuals %>%
  filter(key_name %in% project_keys) %>%
  distinct(device_info_serial, .keep_all = TRUE) %>%
  pull(device_info_serial)

device_info_serials
```

Get detections and write to download directory:

```{r query_detections}
source("functions/download_detections.R")
download_detections(device_info_serials, download_directory, con, FALSE)
```

# Download gull detections

Define dataset properties:

```{r}
project_keys = c("HG_OOSTENDE", "LBBG_ZEEBRUGGE")
download_directory = "../data/interim/gull"
```

Get `device_info_serial`s for that project(s). Note that because of the reuse of tags, there might be less `device_info_serials` than bird individuals:

```{r}
device_info_serials <- individuals %>%
  filter(key_name %in% project_keys) %>%
  distinct(device_info_serial, .keep_all = TRUE) %>%
  pull(device_info_serial)

device_info_serials
```

```{r}
# Take subset for now. TODO: remove
device_info_serials <- device_info_serials[1:4]
```

Get detections and write to download directory:

```{r query_detections}
source("functions/download_detections.R")
download_detections(device_info_serials, download_directory, con, FALSE)
```
