---
title: "Darwin Core mapping"
subtitle: "For: Western Marsh Harriers"
author:
- Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Set locale (so we use UTF-8 character encoding):

```{r}
# This works on Mac OS X, might not work on other OS
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Load libraries:

```{r}
library(tidyverse)
library(magrittr)
library(readr)
library(RPostgreSQL)
library(glue)
```

Set file paths (all paths should be relative to this script):
 
```{r}
individuals_and_track_sessions_file = "../data/interim/wmh/individuals_and_track_sessions.csv"
dwc_occurrence_tagging_file = "../data/processed/wmh/dwc_occurrence_tagging.csv"
dwc_mof_tagging_file = "../data/processed/wmh/dwc_mof_tagging.csv"
```

## Connect to UvA-BiTS database

Get credentials from config file (not included in this repo) and create a database connection:

```{r connect_to_uvabits}
uvabits <- config::get("uvabits")
con = dbConnect(
  dbDriver("PostgreSQL"),
  user = uvabits$username,
  password = uvabits$password,
  host = uvabits$server,
  port = uvabits$port,
  dbname = uvabits$database
)
```

## Set dataset specific settings

```{r}
project_keys <- c("MH_WATERLAND")
datasetID <- "https://doi.org/10.15468/rbguhj"
datasetName <- "Bird tracking - GPS tracking of Western Marsh Harriers breeding near the Belgium-Netherlands border"
institutionCode <- "INBO"
rightsHolder <- "INBO"
license <- "http://creativecommons.org/publicdomain/zero/1.0/"
accessRights <- "http://www.inbo.be/en/norms-for-data-use"
informationWithheld <- "see metadata"
```

# Create occurrence core for bird tagging (human observations)

## Get denormalized individuals and track sessions

To create occurrences for the bird capture/tagging with GPS trackers (= `HumanObservation`) and the associated measurements, we query for all bird individuals in a project and their track session information. This view will be used as a base for the Darwin Core mapping:

```{r}
individuals_and_track_sessions_sql <- glue_sql(
  read_file("../sql/individuals_and_track_sessions.sql"),
  .con = con
)
```

To get an overview of this information, we also run and save the query:

```{r query_individuals_and_track_sessions}
individuals_and_track_sessions <- dbGetQuery(con, individuals_and_track_sessions_sql)
```

Preview results:

```{r}
individuals_and_track_sessions %>% head()
```

Save results:

```{r}
write.csv(individuals_and_track_sessions, file = individuals_and_track_sessions_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```

## Term mapping

Map the source data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml). The mapping is done in SQL:

```{r}
dwc_occurrence_tagging_sql <- glue_sql(
  read_file("../sql/dwc_occurrence_tagging.sql"),
  .con = con
)
```

Show SQL:

```{r}
dwc_occurrence_tagging_sql
```

Run SQL:

```{r query_occurrence_human_obs}
dwc_occurrence_tagging <- dbGetQuery(con, dwc_occurrence_tagging_sql)
```

Preview results:

```{r}
dwc_occurrence_tagging %>% head()
```

Save results:

```{r}
write.csv(dwc_occurrence_tagging, file = dwc_occurrence_tagging_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```

# Create measurement or fact extension for bird tagging

## Term mapping

Map the source data to [Darwin Core Measurement Or Facts](http://rs.gbif.org/extension/dwc/measurements_or_facts.xml). The mapping is done in SQL:

```{r}
dwc_mof_tagging_sql <- glue_sql(
  read_file("../sql/dwc_mof_tagging.sql"),
  .con = con
)
```

Show SQL:

```{r}
dwc_mof_tagging_sql
```

Run SQL:

```{r query_mof_human_obs}
dwc_mof_tagging <- dbGetQuery(con, dwc_mof_tagging_sql)
```

Preview results:

```{r}
dwc_mof_tagging %>% head()
```

Save results:

```{r}
write.csv(dwc_mof_tagging, file = dwc_mof_tagging_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```

# Create occurrence core for bird GPS fixes (machine observations)

```{r}
individuals_and_track_sessions_sql <- glue_sql(
  read_file("sql/individuals_and_track_sessions.sql"),
  .con = con
)
```

