---
title: "Download Digitanimal data in Movebank format"
author: "Peter Desmet"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
```

## Get Digitanimal tag and animal information from spreadsheet

```{r}
metadata <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRQOK3EQ7J9lteAKdyi_fYo_zjw5jUNwpRvcaSI6QxqBDs2q9lkSbEbFAv4jXfzj4teacdiq9AGan-2/pub?gid=0&single=true&output=csv")
metadata_colnames <- colnames(metadata)
```

## Set project

Set project:

```{r}
project_id <- "BE_GRAZERS"
```

## Get reference data (metadata)

For a project, get data on animals, tags and deployments in the [Movebank reference data format](https://www.movebank.org/node/27).

Note: fields indicated with `NOT AVAILABLE` below, are fields that are used in the UvA-BiTS reference data, but that are not applicable here.

Map data:

```{r get_ref_data}
movebank_ref_data <-
  metadata %>%
  filter(project == project_id) %>%
  mutate(
    # animal-comments
    # animal-death-comments
    # animal-earliest-date-born
    # animal-exact-date-of-birth
    `animal-id` = animal_id,
    # animal-latest-date-born
    # animal-nickname: NOT AVAILABLE
    # animal-ring-id: NOT AVAILABLE
    `animal-sex` = recode(sex,
      "F" = "f",
      "M" = "m"
    ),
    `animal-taxon` = scientific_name,
    `animal-taxon-detail` = animal_breed,
    `animal-life-stage` = "adult",
    # animal-mass`: NOT AVAILABLE
    # animal-reproductive-condition
    `attachment-type` = "harness",
    # behavior-according-to
    # data-processing-software
    # deploy-off-latitude
    # deploy-off-longitude
    # deploy-off-person
    `deploy-off-timestamp` = format(
      as.POSIXct(track_session_end_date, tz = "UTC", format = "%Y-%m-%d %H:%M:%S"),
      "%Y-%m-%dT%H:%M:%SZ"
    ),
    # deploy-on-latitude: NOT AVAILABLE
    # deploy-on-longitude: NOT AVAILABLE
    # deploy-on-person
    `deploy-on-timestamp` = format(
      as.POSIXct(track_session_start_date, tz = "UTC", format = "%Y-%m-%d %H:%M:%S"),
      "%Y-%m-%dT%H:%M:%SZ"
    ),
    `deployment-comments` = track_session_remarks,
    # deployment-end-comments
    `deployment-end-type` = case_when(
      str_detect(track_session_remarks, "dead") ~ "dead",
      str_detect(track_session_remarks, "defect") ~ "equipment failure",
      TRUE ~ NA_character_
    ),
    # deployment-id: NOT AVAILABLE
    # duty-cycle
    # geolocator-calibration
    # geolocator-light-threshold
    # geolocator-sensor-comments
    # geolocator-sun-elevation-angle
    # habitat-according-to
    # location-accuracy-comments: TODO, there is no location-error-numerical/vertical-error-numerical in the gps data
    # manipulation-comments
    `manipulation-type` = "confined",
    `study-site` = release_location, # For birds this is indicated in deployment-comments
    `tag-readout-method` = "other wireless",
    # beacon-frequency
    `sensor-type` = "GPS",
    # tag-comments
    # tag-failure-comments
    `tag-id` = tag_number,
    `tag-manufacturer-name` = "Digitanimal",
    `tag-mass` = tag_weight,
    # tag-processing-type
    # tag-production-date
    # tag-serial-no` = serial_number, # TODO
  ) %>%
  select(!one_of(metadata_colnames)) %>%
  arrange(`tag-id`)
```

Save to CSV:

```{r}
write_csv(movebank_ref_data, file = here("data", "processed", project_id, "movebank_ref_data.csv"), na = "")
```

## Get gps data (detections)

TODO

## Get acceleration data

Not implemented yet.
