---
title: "Download Ornitela data in Movebank format"
author: "Peter Desmet"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
```

## Get Ornitela tag and bird information from spreadsheet

```{r}
metadata <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSU838D5jOOcPSyL21iheYdC1hEwYaR88x9qhJNLOZtmaTuK6hv23Vpz-o1b5erXur4Vw2g9Z2mQchh/pub?gid=0&single=true&output=csv")
```

## Set project and some values

Some values cannot be automatically assumed for the Movebank format and have to be defined by the user. See [here](https://www.movebank.org/node/2381#metadata) for information on all reference data terms.

Set values:

```{r}
project_id <- "DELTATRACK"
animal_life_stage <- "adult"
attachment_type <- "harness"
manipulation_type <- "none"
bird_remarks_is_nickname <- TRUE
```

## Get reference data (metadata)

For a project, get data on animals, tags and deployments in the [Movebank reference data format](https://www.movebank.org/node/27).

Filter data:

```{r}
movebank_ref_data <-
  metadata %>%
  filter(project == project_id)
colnames(movebank_ref_data) <- str_c("orn_", colnames(movebank_ref_data)) # Add orn_ to colnames
```

Map data:

```{r get_ref_data}
movebank_ref_data <-
  movebank_ref_data %>%
  mutate(
"animal-comments" = "",
# animal-death-comments
# animal-earliest-date-born
# animal-exact-date-of-birth
"animal-id" = "",
# animal-latest-date-born
"animal-nickname" = "",
"animal-ring-id" = "",
"animal-sex" = "",
"animal-taxon" = "",
# animal-taxon-detail
"animal-life-stage" = "",
"animal-mass" = "",
# animal-reproductive-condition
"attachment-type" = "",
# behavior-according-to
# data-processing-software
# deploy-off-latitude
# deploy-off-longitude
# deploy-off-person
"deploy-off-timestamp" = "",
"deploy-on-latitude" = "",
"deploy-on-longitude" = "",
# deploy-on-person
"deploy-on-timestamp" = "",
"deployment-comments" = "",
# deployment-end-comments
"deployment-end-type" = "",
"deployment-id" = "",
# duty-cycle
# geolocator-calibration
# geolocator-light-threshold
# geolocator-sensor-comments
# geolocator-sun-elevation-angle
# habitat-according-to
"location-accuracy-comments" = "",
# manipulation-comments
"manipulation-type" = "",
"study-site" = "",
"tag-readout-method" = "",
# beacon-frequency
"sensor-type" = "",
# tag-comments
# tag-failure-comments
"tag-id" = "",
"tag-manufacturer-name" = "",
"tag-mass" = "",
# tag-processing-type
# tag-production-date
"tag-serial-no" = "",
)
```

Remove original columns:

```{r}
movebank_ref_data <-
  movebank_ref_data %>%
  select(!starts_with("orn_")) 
```

Save to CSV:

```{r}
write_csv(movebank_ref_data, path = here("data", "processed", project_id, "movebank_ref_data.csv"), na = "")
```

## Get gps data (detections)

Done via live feed.

## Get acceleration data

Done via live feed.
