---
title: "Download Ornitela data in Movebank format"
author: "Peter Desmet"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
```

## Get Ornitela tag and bird information from spreadsheet

```{r}
metadata <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSU838D5jOOcPSyL21iheYdC1hEwYaR88x9qhJNLOZtmaTuK6hv23Vpz-o1b5erXur4Vw2g9Z2mQchh/pub?gid=0&single=true&output=csv")
metadata_colnames <- colnames(metadata)
```

## Set project

Set project:

```{r}
project_id <- "LBBG_JUVENILE"
```

## Get reference data (metadata)

For a project, get data on animals, tags and deployments in the [Movebank reference data format](https://www.movebank.org/node/27).

Note: [UvA-BiTS reference data](../sql/movebank_ref.sql) has some fields that are not available in the Ornitela metadata. These are indicated with `NOT AVAILABLE` below and are not included in `movebank_ref_data.csv`, as their empty values would lead to import errors in Movebank. Since UvA-BiTS has more fields, use `bind_rows(uvabits_project_metadata, ornitela_project_metadata)` (UvA-BiTS first) to have all columns in the right order when you want to combine metadata.

Map data:

```{r get_ref_data}
movebank_ref_data <-
  metadata %>%
  filter(project == project_id) %>%
  mutate(
    # animal-comments: NOT AVAILABLE
    # animal-death-comments
    # animal-earliest-date-born
    # animal-exact-date-of-birth
    `animal-id` = metal_ring,
    # animal-latest-date-born
    `animal-nickname` = case_when(
      str_detect(bird_name, "OT-") ~ NA_character_, # Exclude names that are the default tag name
      TRUE ~ bird_name
    ),
    `animal-ring-id` = colour_ring,
    `animal-sex` = recode(sex,
      "F" = "f",
      "M" = "m"
    ),
    `animal-taxon` = scientific_name,
    # animal-taxon-detail
    `animal-life-stage` = recode(age,
      "A" = "adult",
      "J" = "juvenile"
    ),
    `animal-mass` = bird_weight,
    # animal-reproductive-condition
    `attachment-type` = "harness",
    # behavior-according-to
    # data-processing-software
    # deploy-off-latitude
    # deploy-off-longitude
    # deploy-off-person
    `deploy-off-timestamp` = format(
      as.POSIXct(track_session_end_date, tz = "UTC", format = "%Y-%m-%d %H:%M:%S"),
      "%Y-%m-%dT%H:%M:%SZ"
    ),
    # deploy-on-latitude: NOT AVAILABLE
    # deploy-on-longitude: NOT AVAILABLE
    # deploy-on-person
    `deploy-on-timestamp` = format(
      as.POSIXct(track_session_start_date, tz = "UTC", format = "%Y-%m-%d %H:%M:%S"),
      "%Y-%m-%dT%H:%M:%SZ"
    ),
    `deployment-comments` = case_when(
      !is.na(track_session_remarks) ~ paste(release_location, track_session_remarks, sep = " | "),
      TRUE ~ release_location
    ),
    # deployment-end-comments
    `deployment-end-type` = case_when(
      str_detect(track_session_remarks, "dead") ~ "dead",
      str_detect(track_session_remarks, "defect") ~ "equipment failure",
      TRUE ~ NA_character_
    ),
    # deployment-id: NOT AVAILABLE
    # duty-cycle
    # geolocator-calibration
    # geolocator-light-threshold
    # geolocator-sensor-comments
    # geolocator-sun-elevation-angle
    # habitat-according-to
    # location-accuracy-comments: NOT AVAILABLE, there is no location-error-numerical/vertical-error-numerical in the gps data
    # manipulation-comments
    `manipulation-type` = case_when(
      release_location == "VOC_OST" ~ "manipulated other",
      # Birds from VOC were hatched from egg, placed in a controlled environment and subjected to behavioural studies
      TRUE ~ "none"
    ),
    `study-site` = project_id,
    `tag-readout-method` = "phone network",
    # beacon-frequency
    `sensor-type` = "GPS",
    # tag-comments
    # tag-failure-comments
    `tag-id` = serial_number,
    `tag-manufacturer-name` = "Ornitela",
    `tag-mass` = tag_weight,
    # tag-processing-type
    # tag-production-date
    `tag-serial-no` = serial_number,
  ) %>%
  select(!one_of(metadata_colnames)) %>%
  arrange(`tag-id`)
```

Save to CSV:

```{r}
write_csv(movebank_ref_data, file = here("data", "processed", project_id, "movebank_ref_data.csv"), na = "")
```

## Get gps data (detections)

Done via live feed. Important notes when comparing number of records:

```R
ornitela_export %>%
  # Exclude records with lat/long = 0: these are not fed to Movebank
  filter(satcount != 0) %>%
  # Exclude duplicated timestamps: these are not fed to Movebank
  distinct(device_id, UTC_datetime, .keep_all = TRUE) %>%
  # Use the same cutoff date
  filter(UTC_datetime <= as.POSIXct("YYYY-MM-DD 23:59:59")
  
EQUALS

# Select "Include undeployed locations" when exporting
movebank_export %>%
  # Exclude SMS records
  filter(`orn:transmission-protocol` == "GPRS")
  # Use the same cutoff date
  filter(timestamp <= as.POSIXct("YYYY-MM-DD 23:59:59")
```

## Get acceleration data

Not implemented yet.
