---
title: "Prepare Movebank data for Zenodo upload (make frictionless)"
author: "Peter Desmet"
date: "`r Sys.Date()`"
output:
  html_document
---

This script prepares Movebank bird tracking data for upload to [Zenodo](https://zenodo.org):

1. Manually download data from a Movebank study (see [instructions](https://github.com/inbo/bird-tracking/issues/131)) to `data/processed/<study_acronym>`.
2. Run this script to automatically make a `datapackage.json` file based on the data.
3. Manually upload data to Zenodo.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(magrittr)
library(here)
library(frictionless)
library(movepub) # devtools::install_github("inbo/movepub")
```

## Set project and years

```{r}
project_id <- "O_WESTERSCHELDE"
versioned_doi <- "https://doi.org/10.5281/zenodo.5879096"
start_year <- 2018
end_year <- 2020
```

Create file paths:

```{r}
dir <- here::here("data", "processed", project_id)
ref_data <- here::here(dir, paste0(project_id, "-reference-data.csv"))
gps_data <- here::here(dir, paste0(project_id, "-gps-", start_year:end_year, ".csv.gz"))
acc_data <- here::here(dir, paste0(project_id, "-acceleration-", start_year:end_year, ".csv.gz"))
```

## Create and write Data Package

Create Data Package, add resources and write to disk

```{r}
dataset <-
  create_package() %>%
  append(c(id = versioned_doi), after = 0) %>%
  movepub::add_resource("reference-data", ref_data, keys = TRUE) %>%
  movepub::add_resource("gps", gps_data, keys = TRUE) %>%
  movepub::add_resource("acceleration", acc_data, keys = TRUE)
write_package(dataset, dir)
```

## Validate data package

Make sure to validate the Data Package using [frictionless-py](https://github.com/frictionlessdata/frictionless-py) from the terminal, using:

```
frictionless validate datapackage.json
```
