---
title: "Download UvA-BiTS metadata as Movebank reference data"
author: "Peter Desmet"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
library(glue)
```

# Connect to UvA-BiTS database

Get connection settings from `odbc.ini` (not included in this repo) and connect to database:

```{r connect_to_uvabits}
con <- dbConnect(odbc::odbc(), "uvabits")
```

# Download metadata as Movebank reference data

For a project, get data on animals, tags and deployments in the [Movebank reference data format](https://www.movebank.org/node/27). Some values have to be defined by the user. See [here](https://www.movebank.org/node/2381#metadata) for information on all reference data terms.

Set values:

```{r}
project <- "MH_WATERLAND"
animal_life_stage <- "adult" # https://www.movebank.org/node/2381#animal_life_stage
attachment_type <- "harness" # https://www.movebank.org/node/2381#attachment_type
manipulation_type <- "none"  # https://www.movebank.org/node/2381#manipulation_type
animal_name_in_remarks <- TRUE
```

Query data:

```{r}
sql <- glue_sql(read_file(here("sql", "movebank_reference_data.sql")), .con = con)
movebank_reference_data <- dbGetQuery(con, sql)
```

Save to CSV:

```{r}
write_csv(movebank_reference_data, path = here("data", "processed", project, "movebank_reference_data.csv"), na = "")
```
