---
title: "Prepare (meta)data for Movebank upload"
author: "Sanne Govaert"
date: "`r Sys.Date()`"
output:
  html_document
---

This script prepares [Ornitela](https://www.ornitela.com/), [Interrex](http://interrex-tracking.com/) and [Druid](https://druid.tech/) bird tracking data for upload to [Movebank](https://www.movebank.org/). For a project, it downloads **reference data** from a Google spreadsheet where these data are maintained and transforms these to the [Movebank data format](https://www.movebank.org/node/2381) using `dplyr::mutate()` (see below). **GPS data** are automatically synced with Movebank (see `Live Feeds` in a Movebank study). This is not yet implemented for **acceleration data** which are currently not uploaded.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
```

Set user-defined values:

```{r}
project_id <- "CREX_VLAANDEREN"
```

Load custom function to map the spreadsheet to Movebank reference data:

```{r}
source(here::here("src", "functions", "get_reference_data.R"))
```

## Get reference data

```{r}
movebank_ref_data <- get_reference_data(project_id)
```

Save to CSV:

```{r}
readr::write_csv(movebank_ref_data, file = here::here("data", "processed", project_id, "movebank_ref_data.csv"), na = "")
```

## Get gps data (detections)

Done via live feed.

## Get acceleration data

Not applicable.
