---
title: "Prepare reference data for Movebank upload"
author: "Sanne Govaert"
date: "`r Sys.Date()`"
output:
  html_document
---

This script prepares reference data maintained in a **Google Spreadsheet** for upload to [Movebank](https://www.movebank.org/). It supports the manufacturers:

- [Ornitela](https://www.ornitela.com/)
- [Interrex](http://interrex-tracking.com/)
- [Druid](https://druid.tech/)
- [Global Messenger](https://www.gm-tracking.com/)

The GPS data are expected to be automatically synced with Movebank.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)
library(here)
library(readr)
library(purrr)
```

Set user-defined values:

```{r}
project_id <- "BOP_RODENT"
```

Load custom functions to read and map the spreadsheet to Movebank reference data:

```{r}
source(here::here("src", "functions", "get_reference_data.R"))
```

## Get reference data

Get reference data and map to Movebank format:

```{r}
movebank_ref_data <- get_reference_data(project_id)
```

Save to CSV:

```{r}
readr::write_csv(movebank_ref_data, file = here::here("data", "processed", project_id, "movebank_ref_data.csv"), na = "")
```
